---
title: "Classification and characterization of anthropogenic plant communities at the ecoregion level"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Iberian Atlantic anthropogenic vegetation

Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Greta Ivesdal^2^, Adrián Lázaro-Lobo^1^ (<https://orcid.org/0000-0002-0509-2085>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Víctor González-García: Methodology; Investigation; Formal Analysis; Writing – Review & Editing. Greta Ivesdal: Investigation; Formal Analysis; Writing – Review & Editing. Adrián Lázaro-Lobo: Investigation; Writing – Review & Editing. Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Funding

Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico (SV-23-GIJON-JBA).

# Conflict of interest statement

None declared.

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic> stored at https://zenodo.org/doi/10.5281/zenodo.11613571.

. Upon publication, a version of record of the repository will be deposited in Zenodo.